Highly Available Transactions: Virtues and Limitations\\(Extended Version)

### 2013-07-25 writeup/evaluation.tex
Removing a few references, squeezing space

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failures. (Note,
however, that lock-based protocols \textit{do} offer the benefit of
recency guarantees.) While multi-versioned storage systems allow for a
variety of transactional guarantees, few offer traditional weak
isolation (e.g., non-``tentative update'' schemes) in this context.
The MDCC~\cite{mdcc} protocol offers Read Committed isolation with
Lost Update avoidance but is similarly unavailable due to its reliance
on preventing write conflicts. Chan and Gray's read-only transactions
have item-cut isolation with causal consistency and RA (session
\textit{PL-2L}~\cite{adya}) but are unavailable in the presence of coordinator
failure and assume serializable update transactions~\cite{readonly};
this is similar to read-only and write-only transactions more recently
proposed by Eiger~\cite{eiger}. Causal Serializability offers a
similar model (with unavailable implementation): causal consistency
with a variant of Read Uncommitted between transactions that write to
the same data item~\cite{raynal-causal}.  Brantner's S3
database~\cite{kraska-s3} and Bayou~\cite{sessionguarantees} can all
provide variants of session \textit{PL-2L} with high availability, but none
provide this HAT functionality without substantial
modification. Swift~\cite{swift} and bolt-on causal
consistency~\cite{bolton} are closest to providing maximal sticky HAT
semantics. Accordingly, it is possible to implement many
guarantees weaker than serializability---including guarantees that are
achievable with high availability---and still not achieve high
availability.

>>>>

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failures. (Note, however, that lock-based
protocols \textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, few offer traditional weak isolation (e.g.,
non-``tentative update'' schemes) in this context.  Chan and Gray's
read-only transactions have item-cut isolation with causal consistency
and RA (session \textit{PL-2L}~\cite{adya}) but are unavailable in the
presence of coordinator failure and assume serializable update
transactions~\cite{readonly}; this is similar to read-only and
write-only transactions more recently proposed by
Eiger~\cite{eiger}.  Brantner's S3 database~\cite{kraska-s3}
and Bayou~\cite{sessionguarantees} can all provide variants of session
\textit{PL-2L} with high availability, but none provide this HAT
functionality without substantial modification. Swift~\cite{swift} and
bolt-on causal consistency~\cite{bolton} are closest to providing
maximal sticky HAT semantics. Accordingly, it is possible to implement
many guarantees weaker than serializability---including guarantees
that are achievable with high availability---and still not achieve
high availability.

### 2013-07-25 writeup/hat-analysis.tex
Removing a few references, squeezing space

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, then, for transactional availability,
the server must allow $T_1$ to commit. If the same client subsequently
executes $T_2$ against the same (partitioned) server in the same
session, then it will be able to read its writes. However, if the
network topology changes and the client can only execute $T_2$ on a
different replica that is partitioned from the replica that executed
$T_1$, then the system will have to either stall indefinitely to allow
the client to read her writes (violating transactional availability)
or will have to sacrifice read your writes guarantees. However, if the
client remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, cops, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier: only reveal new writes to
clients when their dependencies have been revealed.

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-1em}
\end{align*}
If we want to guarantee both cut isolation and RA isolation and the
system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs to read
$a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute (e.g.,
COPS~\cite{cops}). Using client-side caching can alleviate some of
these challenges~\cite{bolton, swift} at the expense of requiring
sticky availability.

>>>>

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, then, for transactional availability,
the server must allow $T_1$ to commit. If the same client subsequently
executes $T_2$ against the same (partitioned) server in the same
session, then it will be able to read its writes. However, if the
network topology changes and the client can only execute $T_2$ on a
different replica that is partitioned from the replica that executed
$T_1$, then the system will have to either stall indefinitely to allow
the client to read her writes (violating transactional availability)
or will have to sacrifice read your writes guarantees. However, if the
client remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, eiger, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier: only reveal new writes to
clients when their dependencies have been revealed.

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-1em}
\end{align*}
If we want to guarantee both cut isolation and RA isolation and the
system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs to read
$a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute. Using client-side
caching can alleviate some of these challenges~\cite{bolton, swift} at
the expense of requiring sticky availability.

### 2013-07-25 writeup/intro.tex
Removing a few references, squeezing space

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. Yet, as
formally proven~\cite{gilbert-cap}, the CAP Theorem concerns
linearizability: the ability to read the most recent write to a data
item that is replicated across servers~\cite{herlihy-art}. Yet the
implications of CAP---which concern a fairly narrow form of
distributed data consistency---have often been conflated with the
ability to provide ACID database properties~\cite{hn,brewer-slides,
  foundation-article}; this misunderstanding has led to substantial
confusion regarding replica consistency, transactional isolation, and
high availability. The recent resurgence of transactional systems like
Spanner~\cite{spanner} suggests that programmers value transactional
semantics, but, unfortunately, existing transactional data stores do
not provide availability in the presence of
partitions~\cite{middleware-db, foundation-article,
  hstore,spanner,generalizedsnapshot, mdcc, krikellas-bargain, eiger,
  walter, calvin}.

@@

To date, the relationship between ACID semantics and high
availability has not been well explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, krikellas-bargain, calvin, kemme-classification} or snapshot
isolation~\cite{daudjee-snapshot,generalizedsnapshot, kemme-snapshot,
  walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

>>>>

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. Yet, as
formally proven~\cite{gilbert-cap}, the CAP Theorem concerns
linearizability: the ability to read the most recent write to a data
item that is replicated across servers~\cite{herlihy-art}. Yet the
implications of CAP---which concern a fairly narrow form of
distributed data consistency---have often been conflated with the
ability to provide ACID database
properties~\cite{hat-hotos,brewer-slides, foundation-article}; this
misunderstanding has led to substantial confusion regarding replica
consistency, transactional isolation, and high availability. The
recent resurgence of transactional systems like Spanner~\cite{spanner}
suggests that programmers value transactional semantics, but,
unfortunately, existing transactional data stores do not provide
availability in the presence of partitions~\cite{foundation-article,
  hstore,spanner,eiger,  walter, calvin}.

@@

To date, the relationship between ACID semantics and high
availability has not been well explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, calvin, kemme-classification} or snapshot
isolation~\cite{daudjee-snapshot, walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

### 2013-07-25 writeup/relatedwork.tex
Removing a few references, squeezing space

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  mdcc, eiger, walter,calvin, swift} and industry~\cite{orleans,
  spanner}. As discussed in Section~\ref{sec:modernacid}, classic ACID
databases provide strong semantics but their lock-based and
traditional multi-versioned implementations are unavailable in the
presence of partitions~\cite{bernstein-book, gray-isolation}. Notably,
Google's Spanner provides strong one-copy serializable
transactions. While Spanner is highly specialized for Google's
read-heavy workload, it relies on two-phase commit and two-phase
locking for read/write transactions~\cite{spanner}. As we have
discussed, the latency penalties associated with this design choice
are fundamental to serializability. For users willing to tolerate
these costs, Spanner, or similar strongly consistent, unavailable
systems---including Calvin~\cite{calvin}, G-Store~\cite{gstore},
Generalized Snapshot Isolation~\cite{generalizedsnapshot}, HBase,
HStore~\cite{hstore}, MDCC~\cite{mdcc}, Orleans~\cite{orleans},
Postgres-R~\cite{kemme-thesis}, Walter~\cite{walter}, and a range of
snapshot isolation techniques~\cite{middleware-db, kemme-snapshot,
  daudjee-snapshot}---are a reasonable choice.

>>>>

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  eiger, walter,calvin, swift} and industry~\cite{spanner}. As
discussed in Section~\ref{sec:modernacid}, classic ACID databases
provide strong semantics but their lock-based and traditional
multi-versioned implementations are unavailable in the presence of
partitions~\cite{bernstein-book, gray-isolation}. Notably, Google's
Spanner provides strong one-copy serializable transactions. While
Spanner is highly specialized for Google's read-heavy workload, it
relies on two-phase commit and two-phase locking for read/write
transactions~\cite{spanner}. As we have discussed, the latency
penalties associated with this design choice are fundamental to
serializability. For users willing to tolerate these costs, Spanner,
or similar strongly consistent, unavailable systems---including
Calvin~\cite{calvin}, G-Store~\cite{gstore}, HBase, HStore~\cite{hstore}, Postgres-R~\cite{kemme-thesis},
Walter~\cite{walter}, and a range of snapshot isolation
techniques~\cite{daudjee-snapshot}---are a reasonable choice.
